Beam-Helicity and Beam-Charge Asymmetries\\ Associated with Deeply Virtual Compton Scattering\\ on the Unpolarised Proton \underline{Paper Tag: PDVCS0607 (v3.0)} \underline{Drafting Committee: DC90 (\today)}

### 2012-04-24 submission/paper.tex
Updated some plotting macros and added the submitted paper directory to the repo

Events in the 2006-2007 data set were selected if having exactly one
\red{lepton} track detected within the acceptance of the spectrometer and \red{at
  least one photon depositing $>5$\,GeV in the electromagnetic
  calorimeter. The latter selection criterion differs from the photon
  selection criterion used for the 1996-2005 data set as an intermittent
  hardware fault in 2006-2007 can cause spurious noise signals in the
  calorimeter that are misinterpreted as very low energy photons.}
The event selection is subject to the kinematic constraints 1\,GeV$^{2}$ $<$
Q$^{2}$ $<$ 10\,GeV$^{2}$, 0.03 $<$ $x_{\textrm{B}}$ $<$ 0.35,
$-t < 0.7$\,GeV$^2$, $W^{2}$ $>$
9\,GeV$^{2}$ and $\nu$ $<$ 22\,GeV, where $W$ is the invariant mass of the
$\gamma^{*}p$ system and $\nu$ is the energy of the virtual photon in the target
rest frame. The polar angle between the directions of the virtual and real photons was required to
be within the limits 5~mrad $<$
$\theta_{\gamma^{*}\gamma}$ $<$ 45~mrad.

@@

\section{Background Corrections and Systematic Uncertainties for the
  2006-2007 Data}
The extracted asymmetry amplitudes are subject to systematic uncertainties
that result from a combination of background processes, shifts in the
missing-mass distributions, and various detector and binning effects
determined in the same manner as used in refs.~\cite{Air08,Air09} in order to
maintain consistency with the results published in ref.~\cite{Air09} and
therefore facilitate the combination of the two data sets. \red{No systematic
  uncertainty is assigned from the intermittent fault in the calorimeter
  mentioned in section 2; the percentage of events affected by the issue is
  small for the 2006-2007 data sample and completely negligible in the context
of the combined data sets.}

The contribution to the uncertainties on the amplitude measurements
arising from background in the data from neutral meson
production is predominantly due to the failure to identify one of
  the two photons from the decay of these neutral mesons. \red{The
    $\mathcal{A}^{\textrm{DVCS}}_{\textrm{LU}}$ amplitudes are corrected for
  influences from semi-inclusive pion production and an uncertainty is
  assigned to the affected amplitudes. A further uncertainty is assigned to
  these amplitudes for the influence of photons produced in decays of
  exclusively produced pions.} The procedure for estimating the corresponding
uncertainty on the measured amplitude values is described in detail in
refs.~\cite{Air08,Air09}. Each measurement in the $-t$, $x_{\textrm{B}}$ and
  $Q^{2}$ projections has this uncertainty estimated at the
  centre of the relevant kinematic bin and included as part of the
  total systematic uncertainty. \red{No correction is applied or uncertainty
    assigned to the amplitudes for the
    $\mathcal{A}_{\textrm{LU}}^{\textrm{I}}$ or $\mathcal{A}_{\textrm{C}}$
    asymmetries because influences from meson production are expected to
    vanish when considering the difference between beam charges.}

@@

Table~\ref{table_systematic_contributions_0607} shows, for each
physically-motivated amplitude extracted from the 2006-2007 data (in a
single, overall kinematic bin), the contribution of the various
systematic uncertainties. The beam polarisation measurements have
total uncertainties of 2.8\% for the 1996-2005 data-taking period and
3.4\% for the 2006-2007 data taking period. These uncertainties are
present in the beam-helicity amplitudes and are\red{, as independent
  scale uncertainties, not included} in the other presented uncertainties.
\begin{table}
 \begin{center}
\resizebox{\textwidth}{!} {
 \begin{tabular}{|c|c||c|c|c|c|c|}
  \hline
 & $A$ $\pm$ $\delta_{stat.}$& Background & Missing-Mass Shift  & Detector Effects & & Total $\delta_{syst.}$ \\
  \hline
  \hline
  $A_{\textrm{LU,I}}^{\sin\phi}$ & -0.222  $\pm$  0.023  & 0.002 & 0.001 & 0.022 & & 0.022 \\
  \hline
  $A_{\textrm{LU,DVCS}}^{\sin\phi}$ & 0.005  $\pm$  0.023  & 0.002 & 0.002 & 0.001 & & 0.003 \\
  \hline
  $A_{\textrm{LU,I}}^{\sin(2\phi)}$ & 0.005  $\pm$  0.023  & 0.003 & 0.001 & 0.001 & & 0.003 \\
  \hline
  \hline
  $A_{\textrm{C}}^{\cos(0\phi)}$ & -0.024 $\pm$  0.004 & 0.001 & 0.003 & 0.010 & & 0.011 \\
  \hline
  $A_{\textrm{C}}^{\cos\phi}$ & 0.032  $\pm$  0.006 & 0.002 & 0.001 & 0.001 & & 0.002 \\
  \hline
  $A_{\textrm{C}}^{\cos(2\phi)}$ & -0.004  $\pm$  0.005 & 0.001 & 0.000 & 0.014 & & 0.014 \\
  \hline
  $A_{\textrm{C}}^{\cos(3\phi)}$ & 0.001  $\pm$   0.005 & 0.000 & 0.001 & 0.003 & & 0.004 \\
  \hline
 \end{tabular}
}
  \caption{The values of the physically-motivated asymmetries extracted in a single bin over all kinematic variables with their statistical uncertainties are presented in the second column of this table. The third, fourth and fifth columns show the contributions to the overall systematic uncertainties of the extracted asymmetry amplitudes due to the background correction, the time-dependent shifts of the missing-mass distributions and detector effects for the 2006-2007 data. The total
systematic uncertainties of the amplitudes, shown in the right-most column, are the individual contributions added in quadrature.}
  \label{table_systematic_contributions_0607}
\end{center}
\end{table}
\section{Results}
In figures~\ref{release_bsa_0607} and~\ref{release_bca_0607}, results for the beam-helicity and beam-charge asymmetry amplitudes extracted from the 2006-2007 hydrogen data sets in this work are compared with results extracted from the 1996-2005\footnote{The hydrogen gas target for the data set of 1996-2005 was either unpolarised, transversely polarised or longitudinally polarised. However, the time-averaged polarisation of the polarised targets was negligible, while the rapid reversal time (60--180s) of the polarisation direction minimised bias due to detector effects.} data set published previously~\cite{Air09}. Each of the asymmetry amplitudes is shown extracted in one bin over all kinematic variables (``Overall'') and also projected against $-t$, $x_{\textrm{B}}$ and $Q^{2}$. The beam-helicity asymmetry amplitudes are subject to an additional scale uncertainty from the measurement of the beam polarisation, which is stated in the captions of the figures.
\begin{figure}
\begin{center}
\includegraphics[width=15cm,keepaspectratio]{bsa_sep}
  \caption{Beam-helicity asymmetry amplitudes extracted separately from
the unpolarised 1996-2005 (open triangles)~\cite{Air09} and 2006-2007 (filled squares)
hydrogen data. The error bars represent the statistical uncertainties. The error bands represent the systematic uncertainties.
An additional 2.8\,\% and 3.4\,\% scale uncertainty for the 1996-2005 and
2006-2007 data respectively is present in the amplitudes due to the uncertainty of
the beam polarisation measurement. The simulated fractional contribution from associated production to the yield in each kinematic bin is shown in the bottom row.}
 \label{release_bsa_0607}
\end{center}
 \end{figure}

@@

\section{Summary}
Beam-helicity and beam-charge asymmetries in the azimuthal
distribution of real photons from \red{hard exclusive} leptoproduction on an unpolarised
hydrogen target have been presented. These asymmetries were extracted
from an unpolarised hydrogen data set taken during the 2006 and 2007
operating periods of H{\sc ermes}. Analogous asymmetry amplitudes were
extracted previously from hydrogen data obtained during the 1996-2005
experimental period as described in ref.~\cite{Air09}. A comparison of
the amplitudes extracted from these independent data sets has shown
that they are compatible and the asymmetry amplitudes can therefore be
extracted from the complete 1996-2007 data set. The asymmetry
amplitudes extracted from the complete data set are the most
statistically precise DVCS measurements presented by H{\sc
  ermes}. There is a strong signal in the first harmonic of the
interference contribution to the beam-helicity asymmetry. There are
non-zero amplitudes in the zeroth and first harmonics of the
beam-charge asymmetry. All asymmetry amplitudes related to higher
harmonics are consistent with zero. The results from the complete data
set are compared to calculations from ongoing work to fit GPD models to experimental data. All asymmetry amplitudes are also presented as projections in $-t$ in bins of $x_{\textrm{B}}$. No additional features are observed in any particular $x_{\textrm{B}}$-bin.

@@

  single bin across the whole kinematic range. \red{The size of the symbols in

@@

(open) symbols represent positive (negative) correlations.}}

>>>>

Events in the 2006-2007 data set were selected if having exactly one
lepton track detected within the acceptance of the spectrometer and at
  least one photon depositing $>5$\,GeV in the electromagnetic
  calorimeter. The latter selection criterion differs from the photon
  selection criterion used for the 1996-2005 data set as an intermittent
  hardware fault in 2006-2007 can cause spurious noise signals in the
  calorimeter that are misinterpreted as very low energy photons.
The event selection is subject to the kinematic constraints 1\,GeV$^{2}$ $<$
Q$^{2}$ $<$ 10\,GeV$^{2}$, 0.03 $<$ $x_{\textrm{B}}$ $<$ 0.35,
$-t < 0.7$\,GeV$^2$, $W^{2}$ $>$
9\,GeV$^{2}$ and $\nu$ $<$ 22\,GeV, where $W$ is the invariant mass of the
$\gamma^{*}p$ system and $\nu$ is the energy of the virtual photon in the target
rest frame. The polar angle between the directions of the virtual and real photons was required to
be within the limits 5~mrad $<$
$\theta_{\gamma^{*}\gamma}$ $<$ 45~mrad.

@@

\section{Background Corrections and Systematic Uncertainties for the
  2006-2007 Data}
The extracted asymmetry amplitudes are subject to systematic uncertainties
that result from a combination of background processes, shifts in the
missing-mass distributions, and various detector and binning effects
determined in the same manner as used in refs.~\cite{Air08,Air09} in order to
maintain consistency with the results published in ref.~\cite{Air09} and
therefore facilitate the combination of the two data sets. No systematic
  uncertainty is assigned from the intermittent fault in the calorimeter
  mentioned in section 2; the number of events in which the fault
  occurred is very small for the 2006-2007 data sample and completely negligible in the context
of the combined data sets.

The contribution to the uncertainties on the amplitude measurements
arising from background in the data from neutral meson
production is predominantly due to the failure to identify one of
  the two photons from the decay of these neutral mesons. The
    $\mathcal{A}^{\textrm{DVCS}}_{\textrm{LU}}$ amplitudes are corrected for
  the fraction of the data sample and the magnitude of the asymmetry
  due to semi-inclusive pion production. A further uncertainty is assigned to
  these amplitudes for the influence of photons produced in decays of
  exclusively produced pions. Corrections for dilution of the amplitudes for the
    $\mathcal{A}_{\textrm{LU}}^{\textrm{I}}$ and $\mathcal{A}_{\textrm{C}}$
    asymmetries are applied. No asymmetry value is assigned to the
    influence of the dilution because influences from meson production are expected to
    vanish when considering the difference between beam charges.
The procedure for estimating the
uncertainty and correction factor for each measured amplitude value is described in detail in
refs.~\cite{Air08,Air09}. Each measurement in the $-t$, $x_{\textrm{B}}$ and
  $Q^{2}$ projections has this uncertainty estimated at the
  centre of the relevant kinematic bin and included as part of the
  total systematic uncertainty.

@@

Table~\ref{table_systematic_contributions_0607} shows, for each
physically-motivated amplitude extracted from the 2006-2007 data (in a
single, overall kinematic bin), the contribution of the various
systematic uncertainties. The beam polarisation measurements have
total uncertainties of 2.8\% for the 1996-2005 data-taking period and
3.4\% for the 2006-2007 data taking period. These uncertainties are
present in the beam-helicity amplitudes and are, as independent
  scale uncertainties, not included in the other presented uncertainties.
\begin{table}
 \begin{center}
\resizebox{\textwidth}{!} {
 \begin{tabular}{|c|c||c|c|c|c|c|}
  \hline
 & $A$ $\pm$ $\delta_{stat.}$& Background & Missing-Mass Shift  & Detector Effects & & Total $\delta_{syst.}$ \\
  \hline
  \hline
  $A_{\textrm{LU,I}}^{\sin\phi}$ & -0.222  $\pm$  0.023  & 0.002 & 0.001 & 0.022 & & 0.022 \\
  \hline
  $A_{\textrm{LU,DVCS}}^{\sin\phi}$ & 0.005  $\pm$  0.023  & 0.002 & 0.002 & 0.001 & & 0.003 \\
  \hline
  $A_{\textrm{LU,I}}^{\sin(2\phi)}$ & 0.005  $\pm$  0.023  & 0.003 & 0.001 & 0.001 & & 0.003 \\
  \hline
  \hline
  $A_{\textrm{C}}^{\cos(0\phi)}$ & -0.024 $\pm$  0.004 & 0.001 & 0.003 & 0.010 & & 0.011 \\
  \hline
  $A_{\textrm{C}}^{\cos\phi}$ & 0.032  $\pm$  0.006 & 0.002 & 0.001 & 0.001 & & 0.002 \\
  \hline
  $A_{\textrm{C}}^{\cos(2\phi)}$ & -0.004  $\pm$  0.005 & 0.001 & 0.000 & 0.014 & & 0.014 \\
  \hline
  $A_{\textrm{C}}^{\cos(3\phi)}$ & 0.001  $\pm$   0.005 & 0.000 & 0.001 & 0.003 & & 0.004 \\
  \hline
 \end{tabular}
}
  \caption{The values of the physically-motivated asymmetries extracted in a single bin over all kinematic variables with their statistical uncertainties are presented in the second column of this table. The third, fourth and fifth columns show the contributions to the overall systematic uncertainties of the extracted asymmetry amplitudes due to the background correction, the time-dependent shifts of the missing-mass distributions and detector effects for the 2006-2007 data. The total
systematic uncertainties of the amplitudes, shown in the right-most column, are the individual contributions added in quadrature.}
  \label{table_systematic_contributions_0607}
\end{center}
\end{table}
\section{Results}
In figures~\ref{release_bsa_0607} and~\ref{release_bca_0607}, results for the beam-helicity and beam-charge asymmetry amplitudes extracted from the 2006-2007 hydrogen data sets in this work are compared with results extracted from the 1996-2005\footnote{The hydrogen gas target for the data set of 1996-2005 was either unpolarised, transversely polarised or longitudinally polarised. However, the time-averaged polarisation of the polarised targets was negligible, while the rapid reversal time (60--180s) of the polarisation direction minimised bias due to detector effects.} data set published previously~\cite{Air09}. Each of the asymmetry amplitudes is shown extracted in one bin over all kinematic variables (``Overall'') and also projected against $-t$, $x_{\textrm{B}}$ and $Q^{2}$. The beam-helicity asymmetry amplitudes are subject to an additional scale uncertainty from the measurement of the beam polarisation, which is stated in the captions of the figures.
\begin{figure}
\begin{center}
\includegraphics[width=15cm,keepaspectratio]{bsa_sep}
  \caption{Beam-helicity asymmetry amplitudes extracted separately from
the unpolarised 1996-2005 (open triangles)~\cite{Air09} and 2006-2007 (filled squares)
hydrogen data. The error bars represent the statistical uncertainties. The error bands represent the systematic uncertainties.
An additional 2.8\,\% and 3.4\,\% scale uncertainty for the 1996-2005 and
2006-2007 data respectively is present in the amplitudes due to the uncertainty of
the beam polarisation measurement. The simulated fractional contribution from associated production to the yield in each kinematic bin is shown in the bottom row.}
 \label{release_bsa_0607}
\end{center}
 \end{figure}

@@

\section{Summary}
Beam-helicity and beam-charge asymmetries in the azimuthal
distribution of real photons from hard exclusive leptoproduction on an unpolarised
hydrogen target have been presented. These asymmetries were extracted
from an unpolarised hydrogen data set taken during the 2006 and 2007
operating periods of H{\sc ermes}. Analogous asymmetry amplitudes were
extracted previously from hydrogen data obtained during the 1996-2005
experimental period as described in ref.~\cite{Air09}. A comparison of
the amplitudes extracted from these independent data sets has shown
that they are compatible and the asymmetry amplitudes can therefore be
extracted from the complete 1996-2007 data set. The asymmetry
amplitudes extracted from the complete data set are the most
statistically precise DVCS measurements presented by H{\sc
  ermes}. There is a strong signal in the first harmonic of the
interference contribution to the beam-helicity asymmetry. There are
non-zero amplitudes in the zeroth and first harmonics of the
beam-charge asymmetry. All asymmetry amplitudes related to higher
harmonics are consistent with zero. The results from the complete data
set are compared to calculations from ongoing work to fit GPD models to experimental data. All asymmetry amplitudes are also presented as projections in $-t$ in bins of $x_{\textrm{B}}$. No additional features are observed in any particular $x_{\textrm{B}}$-bin.

@@

  single bin across the whole kinematic range. The size of the symbols in

@@

(open) symbols represent positive (negative) correlations.}
